---
title: "Stat 427 Consulting Project Group 7"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---

<style>
body {
text-align: justify}
</style>

```{css, echo=FALSE}
.solution {
background-color: #CCDDFF;
}
```

# Loading the Packages and Data

```{r}
Sys.setenv(LANGUAGE = "en")

# Load Exploratory data analysis packages
library(dlookr)
library(GGally)
library(readxl)
library(tidyverse)

# Load data modeling packages
library(lme4)
library(emmeans)
library(optimx)
library(psych)
library(car)
library(MuMIn)
library(pwr)

# Load data visualization packages
library(ggplot2)
library(ggpubr)

# Load data
priming_dataset <- read_excel("priming_dataset_cleaned_v2.xlsx")

# Remove gender, age, bilingual_type and lang_variety columns
priming_dataset <- priming_dataset %>%
  select(-c(gender,age,bilingual_type,lang_variety))
```

Feature Types:

Numerical Features:

- BLP

- language_use_span

- language_use_eng

- MLU_spa

- Words_Min_spa

- VOCD_spa

- MLU_eng

- Words_Min_eng

- VOCD_eng

Categorical Features:

- subject

- group

- phase

- construction

- mode

- target

- n_item

```{r}
# Convert Categorical features into factors
Cat_features <- c("subject","group","phase","construction","mode","target","n_item")
priming_dataset[,Cat_features] <- lapply(priming_dataset[,Cat_features] , factor)

# View the data
priming_dataset %>%
  head()
```
# Missing Value

```{r}
# Check columns containing missing value
priming_dataset %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) 
```

No missing value

# RQ1

Is the priming effect stronger with the ACC or with the SPE construction?

## Exploratory Data Analysis (Categorical Features)

```{r}
# Filter the data set for RQ1
priming_dataset_rq1 <- priming_dataset %>%
  filter(mode=="within")

# Check the structure of RQ1 data
str(priming_dataset_rq1[,Cat_features])
```

### Explore the relationship between "group" and "target"

```{r}
# Create a contingency table of the "target" and "group"
addmargins(table(priming_dataset_rq1$group, priming_dataset_rq1$target))
```

```{r}
# Create a proportional contingency table of the "target" and "group"
prop.table(table(priming_dataset_rq1$group, priming_dataset_rq1$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq1) +
  aes(x = group, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Group") + 
  ylab("Proportion")
```

```{r}
# Chi-Square concept: https://data-flair.training/blogs/chi-square-test-in-r/

# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq1_sample <- priming_dataset_rq1[sample(length(priming_dataset_rq1$subject), 1000), ]
chisq.test(priming_dataset_rq1_sample$group, priming_dataset_rq1_sample$target)
```

### Explore the relationship between "phase" and "target"

```{r}
# Create a contingency table of the "target" and "phase"
addmargins(table(priming_dataset_rq1$phase, priming_dataset_rq1$target))
```

```{r}
# Create a proportional contingency table of the "target" and "phase"
prop.table(table(priming_dataset_rq1$phase, priming_dataset_rq1$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq1) +
  aes(x = phase, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Phase") + 
  ylab("Proportion")
```

```{r}
# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq1_sample <- priming_dataset_rq1[sample(length(priming_dataset_rq1$subject), 1000), ]
chisq.test(priming_dataset_rq1_sample$phase, priming_dataset_rq1_sample$target)
```

### Explore the relationship between "construction" and "target"

```{r}
# Create a contingency table of the "target" and "construction"
addmargins(table(priming_dataset_rq1$construction, priming_dataset_rq1$target))
```

```{r}
# Create a proportional contingency table of the "target" and "construction"
prop.table(table(priming_dataset_rq1$construction, priming_dataset_rq1$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq1) +
  aes(x = construction, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Construction") + 
  ylab("Proportion")
```

```{r}
# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq1_sample <- priming_dataset_rq1[sample(length(priming_dataset_rq1$subject), 1000), ]
chisq.test(priming_dataset_rq1$construction, priming_dataset_rq1$target)
```

## Data Modeling
### Generalized linear mixed model fit by maximum likelihood - item level

```{r}
# Reorder the level of the features
priming_dataset_rq1$phase = relevel(priming_dataset_rq1$phase, ref = "pre-test")
priming_dataset_rq1$construction = relevel(priming_dataset_rq1$construction, ref = "spe")
priming_dataset_rq1$target = relevel(priming_dataset_rq1$target, ref = "no")

# How to avoid the convergence problem (Change a optimizer):  https://stats.stackexchange.com/questions/242109/model-failed-to-converge-warning-in-lmer

# Run the glm model
glm_rq1 = lme4::glmer(target ~ phase * construction + group + (1|subject)+ (1|n_item),
                     data = priming_dataset_rq1, family = "binomial",
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
summary(glm_rq1)
```


```{r}
# Estimate marginal means by construction
pairs(emmeans(glm_rq1, "phase", by = "construction"))
```


```{r}
# Estimate marginal means by phase
pairs(emmeans(glm_rq1, "construction", by = "phase"))
```

## Power Analysis

```{r}
# Pseudo-R-square for the GLM Function: https://search.r-project.org/CRAN/refmans/MuMIn/html/r.squaredGLMM.html

# Calculate Pseudo-R-squared for Generalized Mixed-Effect models
MuMIn::r.squaredGLMM(glm_rq1)
```

```{r}
# Power Analysis in linguistic area (Literature Review): https://www.jstor.org/stable/3587103

# Use R to calculate the power https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html

# Calculate the power for RQ1
pwr.f2.test(u = 7, v =  124-7-1, f2 = 0.3294668/(1-0.3294668), sig.level = 0.05)
```



# RQ2
Is the priming effect stronger in within-language mode or in cross-linguistic mode?

## Exploratory Data Analysis (Categorical Features)

```{r}
# Filter the data set for RQ1
priming_dataset_rq2 <- priming_dataset %>%
  filter(construction=="spe")

# Check the structure of RQ1 data
str(priming_dataset_rq2[,Cat_features])
```

### Explore the relationship between "group" and "target"

```{r}
# Create a contingency table of the "target" and "group"
addmargins(table(priming_dataset_rq2$group, priming_dataset_rq2$target))
```

```{r}
# Create a proportional contingency table of the "target" and "group"
prop.table(table(priming_dataset_rq2$group, priming_dataset_rq2$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq2) +
  aes(x = group, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Group") + 
  ylab("Proportion")
```

```{r}
# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq2_sample <- priming_dataset_rq2[sample(length(priming_dataset_rq2$subject), 1000), ]
chisq.test(priming_dataset_rq2_sample$group, priming_dataset_rq2_sample$target)
```

### Explore the relationship between "phase" and "target"

```{r}
# Create a contingency table of the "target" and "phase"
addmargins(table(priming_dataset_rq2$phase, priming_dataset_rq2$target))
```

```{r}
# Create a proportional contingency table of the "target" and "phase"
prop.table(table(priming_dataset_rq2$phase, priming_dataset_rq2$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq2) +
  aes(x = phase, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Phase") + 
  ylab("Proportion")
```

```{r}
# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq2_sample <- priming_dataset_rq2[sample(length(priming_dataset_rq2$subject), 1000), ]
chisq.test(priming_dataset_rq2_sample$phase, priming_dataset_rq2_sample$target)
```

### Explore the relationship between "construction" and "target"

```{r}
# Create a contingency table of the "target" and "mode"
addmargins(table(priming_dataset_rq2$mode, priming_dataset_rq2$target))
```

```{r}
# Create a proportional contingency table of the "target" and "mode"
prop.table(table(priming_dataset_rq2$mode, priming_dataset_rq2$target), margin=1)*100
```

```{r}
# Visualize the proportional contingency table
ggplot(priming_dataset_rq2) +
  aes(x = mode, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Mode") + 
  ylab("Proportion")
```

```{r}
# Perform Chi-Square test
set.seed(1) 
priming_dataset_rq2_sample <- priming_dataset_rq2[sample(length(priming_dataset_rq2$subject), 1000), ]
chisq.test(priming_dataset_rq2_sample$mode, priming_dataset_rq2_sample$target)
```

## Data Modeling

### Generalized linear mixed model fit by maximum likelihood - item level

```{r}
# Reorder the level of the features
priming_dataset_rq2$phase = relevel(priming_dataset_rq2$phase, ref = "pre-test")
priming_dataset_rq2$mode = relevel(priming_dataset_rq2$mode, ref = "within")
priming_dataset_rq2$target = relevel(priming_dataset_rq2$target, ref = "no")

# Run the glm model
glm_rq2 = lme4::glmer(target ~ phase * mode + group + (1|subject)+ (1|n_item),
                     data = priming_dataset_rq2, family = "binomial",
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

summary(glm_rq2)
```


```{r}
# Estimate marginal means by mode
pairs(emmeans(glm_rq2, "phase", by = "mode"))
```


```{r}
# Estimate marginal means by phase
pairs(emmeans(glm_rq2, "mode", by = "phase"))
```

### Why mode not significant?

```{r}
# Split the data set based on phase
priming_dataset_rq2_pre <- priming_dataset_rq2 %>%
  filter(phase=="pre-test")
priming_dataset_rq2_treat <- priming_dataset_rq2 %>%
  filter(phase=="treatment")
priming_dataset_rq2_post <- priming_dataset_rq2 %>%
  filter(phase=="post-test")

# Visualize the proportional contingency tables by different phases
pre_plot <- ggplot(priming_dataset_rq2_pre) +
  aes(x = mode, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Mode") + 
  ylab("Proportion")

treat_plot <- ggplot(priming_dataset_rq2_treat) +
  aes(x = mode, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Mode") + 
  ylab("Proportion")

post_plot <- ggplot(priming_dataset_rq2_post) +
  aes(x = mode, fill = target) +
  geom_bar(position = "fill") + 
  xlab("Mode") + 
  ylab("Proportion")

ggarrange(pre_plot, treat_plot, post_plot, 
          labels = c("pre-test", "treatment", "post-test"),
          ncol = 2, nrow = 2)
```

## Power Analysis

```{r}
# Calculate Pseudo-R-squared for Generalized Mixed-Effect models
MuMIn::r.squaredGLMM(glm_rq2)
```

```{r}
# Calculate the power for RQ2
pwr.f2.test(u = 7, v =  124-7-1, f2 = 0.1616664/(1-0.1616664), sig.level = 0.05)
```

# RQ3

Which individual variables are associated with a strong priming effect?

## Exploratory Data Analysis (Numerical Features)

```{r}
# Select numerical features
priming_dataset_rq3 <- priming_dataset %>%
  select(c("subject","n_item","target","BLP","language_use_span","language_use_eng","MLU_spa","Words_Min_spa","VOCD_spa","MLU_eng","Words_Min_eng","VOCD_eng"))

priming_dataset_rq3 %>%
  head()
```

### Pairplot Analysis

```{r}

# Pairplot theme https://ggplot2.tidyverse.org/reference/theme.html
# Pairplot text size: https://stackoverflow.com/questions/8599685/how-to-change-correlation-text-size-in-ggpairs

# Check the pair plot
ggpairs(priming_dataset_rq3, columns = 3:12,upper=list(continuous = wrap("cor",size=2)), aes(colour=target,alpha = 0.5))+theme( strip.text = element_text(size = 5),axis.text = element_text(size = 5))
```

### Correlation analysis

```{r}
# Check strongly correlated variables
cor_matrix <- as.data.frame(cor(priming_dataset_rq3[,4:12],method="pearson"))
cor_matrix[abs(cor_matrix) < 0.5] <- ""
cor_matrix
```

- "BLP" is strongly correlated with "language_use_span","language_use_eng" and "Words_Min_eng"

- "language_use_span" is exactly correlated with "language_use_eng" and strongly correlated with "Words_Min_eng"

```{r}
# Remove "language_use_eng"
priming_dataset_rq3 <- priming_dataset_rq3 %>%
  select(-language_use_eng)

priming_dataset_rq3 %>%
  head()
```

```{r}
# The the mean difference of individual variables between target yes and target no
priming_dataset_rq3 %>%
  group_by(target) %>%
  summarize(BLP_mean=mean(BLP), language_use_span_mean=mean(language_use_span), MLU_spa_mean=mean(MLU_spa), Words_Min_spa_mean=mean(Words_Min_spa), VOCD_spa_mean=mean(VOCD_spa), MLU_eng_mean=mean(MLU_eng), Words_Min_eng_mean=mean(Words_Min_eng), VOCD_eng_mean=mean(VOCD_eng))
```

## Data Modeling 

### Linear Regression (subject level)

We recommend our client to use glm method not linear regression in RQ3. Therefore, this part can be used just as FYI

```{r}
# Simplify the data set into subject level
priming_dataset_rq3_sbj <- priming_dataset_rq3 %>%
  select(-n_item) %>%
  mutate(target_num=ifelse(target=="no",0,1)) %>%
  group_by(subject,BLP,language_use_span,MLU_spa,Words_Min_spa,VOCD_spa,MLU_eng,Words_Min_eng,VOCD_eng) %>%
  summarize(target_mean=mean(target_num)) %>%
  as.data.frame()

priming_dataset_rq3_sbj%>%
  head()
```

```{r}
# Check the pair plot
ggpairs(priming_dataset_rq3_sbj, columns = 2:10)
```

```{r}
# Perform the linear regression model
lm.fit <- lm(target_mean~BLP+language_use_span+MLU_spa+Words_Min_spa+VOCD_spa+MLU_eng+Words_Min_eng+VOCD_eng, data = priming_dataset_rq3_sbj)
summary(lm.fit)
```
```{r}
# vif and multicollinearity https://www.analyticsvidhya.com/blog/2020/03/what-is-multicollinearity/
vif(lm.fit)
```

BLP and language_use_span have vif greater than 10. Therefore, remove one of them (language_use_span) to address the multicollinearity problem.

```{r}
# Perform the linear regression model
lm.fit <- lm(target_mean~BLP+MLU_spa+Words_Min_spa+VOCD_spa+MLU_eng+Words_Min_eng+VOCD_eng, data = priming_dataset_rq3_sbj)
summary(lm.fit)
```

```{r}
# Perform stepwise regression for variable selection
full <- lm(target_mean~BLP+MLU_spa+Words_Min_spa+VOCD_spa+MLU_eng+Words_Min_eng+VOCD_eng, data = priming_dataset_rq3_sbj)

null <- lm(target_mean~1, data = priming_dataset_rq3_sbj)

step(null,scope=list(upper=full,lower=null), data =priming_dataset_rq3_sbj, direction="both")
```

```{r}
# Perform the linear regression model after stepwise variable selection
lm.fit_sig <- lm(target_mean ~ BLP, data = priming_dataset_rq3_sbj)
summary(lm.fit_sig)
```

### Generalized linear mixed model fit by maximum likelihood

```{r}
# Parameters or bounds appear to have different scalings can cause poor performance in optimization. Therefore, scale all the features

# Scale the features
priming_dataset_rq3_std <- priming_dataset_rq3 %>% 
  mutate_at(colnames(priming_dataset_rq3)[4:11], ~(scale(.) %>% as.vector))

# Perform the generalized linear mixed model fit by maximum likelihood
glm_rq3 <- lme4::glmer(target ~ BLP +  language_use_span + MLU_spa + Words_Min_spa + VOCD_spa + MLU_eng + Words_Min_eng + VOCD_eng + (1|subject) + (1|n_item),
                     data = priming_dataset_rq3_std, family = "binomial",
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
summary(glm_rq3)
```

```{r}
# Check the vif score
vif(glm_rq3)
```

```{r}
# Perform the generalized linear mixed model fit by maximum likelihood again after remove language_use_span
glm_rq3 <- lme4::glmer(target ~ BLP + MLU_spa + Words_Min_spa + VOCD_spa + 
                        MLU_eng + Words_Min_eng + VOCD_eng + (1|subject) + (1|n_item),
                     data = priming_dataset_rq3_std, family = "binomial",
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
summary(glm_rq3)
```

```{r}
# Check the vif score again
vif(glm_rq3)
```

```{r}
# Check the model after stepwise variable selection as well
glm_rq3_sig <- lme4::glmer(target ~ BLP + (1|subject) + (1|n_item),
                     data = priming_dataset_rq3_std, family = "binomial",
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
summary(glm_rq3_sig)
```

## Power Analysis

```{r}
# Calculate Pseudo-R-squared for Generalized Mixed-Effect models
MuMIn::r.squaredGLMM(glm_rq3)
```
```{r}
# Calculate the power for RQ3
pwr.f2.test(u = 7, v =  124-7-1, f2 = 0.005760835/(1-0.005760835), sig.level = 0.05)
```

